---
title: "writeup one"
author: "Tian Tan"
date: "April 27, 2016"
output: pdf_document
---

A. Show the probability density of the sum forms a symmetrix triangle above interval(0,2) and derive the cumulative distribution function of X.

1)Proof will use a convoluton.

Let $f_{U_1}$ be the density function of $U_1$, and $f_{U_1}$ be the density function of $U_2$. 

X=$U_1+U_2$. Then
\begin{eqnarray*}
f_X(x) &=& \int_{-\infty}^{\infty} f_{U_1}(u_1)f_{U_2}(x-u_1)du_1
\end{eqnarray*}

Since $U_1$ and $U_2$ are independent uniform random variables on (0,1), the sum, X=$U_1+U_2$, should be in the interval (0,2), and $f_X(x)$=0 for $x\geq 2$ and for x<0.

For $0\leq x<2$, 

i) $0\leq x \leq 1$

$f_{U_1}(u_1)f_{U_2}(x-u_1)$ can be either 1 or 0, 
We need to make sure when $f_{U_2}(x-u_1)$=1, $x-u_1 \geq 0$, that is, $u_1 \leq x$. So we integrate from $u_1$=0 to $u_1$=x.
\begin{eqnarray*}
\int_{0}^{x} 1du_1 &=& x
\end{eqnarray*}
So $f_X(x)=x$ for $0<x<1$

ii) 1<x<2
$f_{U_1}(u_1)f_{U_2}(x-u_1)$ can be either 1 or 0, 
We need to make sure when $f_{U_2}(x-u_1)$=1, $x-u_1 \leq 1$, that is, $u_1 \geq x-1$. So we integrate from $u_1$=x-1 to $u_1$=1.
\begin{eqnarray*}
\int_{x-1}^{2} 1du_1 &=& 2-x
\end{eqnarray*}
So $f_X(x)=2-x$ for $1<x<2$

So the probability density of the sum, X=$U_1+U_2$ forms a symmetric triangle above the interval (0,2)

2)derive the cumulative distribution function of X

i) $0\leq x \leq 1$
\begin{eqnarray*}
F(x) &=& \int_{0}^{x} xdx = \frac{1}{2}x^2
\end{eqnarray*}

ii) 1<x<2
\begin{eqnarray*}
F(x) &=& \frac{1}{2}+\frac{1}{2}-(\int_{x}^{2} (2-x)dx)\\
     &=& -\frac{1}{2}x^2+2x-1
\end{eqnarray*}

B. Write a function to generate values of X using the inverse CDF method. 
```{r}
Finv<-function(u){
  # u is a sequence of random variables
  x=rep(length(u))
  for (i in 1:length(u))
   if (u[i]<=1/2){
     x[i]=sqrt(2*u[i])
   } else{
     x[i]=2-sqrt(2-2*u[i])
   }
   return(x)
}
```

C.Generate a random sample of N=1000 from the triangular distribution and display a histogram of the sample values using break points at intervals of 0.2. 
```{r}
u=runif(1000)
x=Finv(u)
hist(x,breaks=seq(0,2,by=0.2),main="histogram of triangular distribution sample values")
```

D.Tabulate the observed frequency counts in the 10 bins and the corresponding expected values. Use the Pearson chi-squared statistic to assess the agreement between observed and expected counts.
```{r,eval=FALSE}
result=matrix(0,nrow=2,ncol=10)
rownames(result)=c("observed frequency counts","expected values")
id=order(x)
x=x[id]
for (i in 1:10){
  for (k in 1:1000){
    if ((0.2*(i-1)<=x[k]) && (x[k]<=0.2*i)){
         result[1,i]=result[1,i]+1
    }
  }
    if (i<=5){
         result[2,i]=result[2,i]+((0.2*i)^2*0.5-(0.2*(i-1))^2*0.5)*1000
      } else{
         result[2,i]=result[2,i]+(0.4-((0.2*i)^2*0.5-(0.2*(i-1))^2*0.5))*1000
      }
}

save(result,file="~/Documents/Cornell/Statistical computing/HW6/result.rData")

```

```{r}
load(file="~/Documents/Cornell/Statistical computing/HW6/result.rData")
result
chisq.test(result)
```

The chi square p value is greater than 0.05, which means that we fail to reject the null hypothesis that the observed counts is independent of the expected counts. 



Cite: work with Rui Chen, rc687
