---
title: "writeup two"
author: "Tian Tan"
date: "April 29, 2016"
output: pdf_document
---

A.Estimate I using Monte-Carlo integration. Include both the estimate and code in the writeup.

By using the Monte-Carlo method
```{r}
f<-function(x){
  sqrt(1-x^2)
}

MonteCarloIntegral<-function(ftn,a,b,n){
  # ftn is a function of one variable
  # [a,b] is the range of integration
  # n is the sample size  

  I=0
  for (i in 1:n){
     u=runif(n,a,b)
     I[i]=mean(ftn(u))*(b-a)
  }
  return(I)
}

IMC=MonteCarloIntegral(f,0,1,1000)
# The estimated value is 
mean(IMC)
```


B. Estimate I using antithetic sampling, and compute an estimate of the percentage variance reduction achieved by using the antithetic approach. Include the estimates and code in the writeup.
```{r}
AntitheticSampling<-function(ftn,n,a,b){
  # ftn is a function of one variable 
  # [a,b] is the range of integration
  # n is the sample size 
  I=0
  for (i in 1:n){
      ua=runif(n,a,b)
      I[i]=0.5*mean(ftn(ua))+0.5*mean(ftn(1-ua))
  }
  return(I)
}
 
IAS=AntitheticSampling(f,1000,0,1)
# The estimated value is
mean(IAS)


# The percentage variance reduction is
(var(IMC)-var(IAS))/var(IMC) *100
```


C. Approximate the integrand by a straight line and use a control variate approach to estimate the value of the integral. Estimate the resulting variance reduction achieved. Specify the control variate used and include the estimate and code in the writeup.
```{r}
ControlCovariate<-function(ftn,n){
  estimator=rep(0,n)
  for (i in 1:n){
    x = runif(n)
    fx = ftn(x)  # Integral we want to estimate
    hx = 2*x+1    # Integral that we know
    alpha = cov(fx,hx)/var(hx)    # Correction factor
    estimator[i]=mean(fx)-alpha*(mean(hx)-2)
  }
  return(list(I=mean(estimator),Variance=var(estimator)))
}
result=ControlCovariate(f,1000)

# The estimated value 
result$I

# The variance reduction
var(IMC)-result$Variance
```


D. Use importance sampling to estimate the integral I. Try using three importance sampling densities, and compare their effectiveness. Specify these importance sampling densities, estimates, and include code in the writeup.

I choose distributions as such: 

X1~U(0,1), density=1

X2~E(6), density= $6e^{-6x}$

X3~beta(1,2), density=2*(1-x)

```{r,eval=FALSE}
ImportanceSampling<-function(ftn,n){
  # ftn is a function of one variable 
  # [a,b] is the range of integration
  # n is the sample size 
  Est1=rep(0,n)
  Est2=rep(0,n)
  Est3=rep(0,n)
  result=matrix(0,nrow=2,ncol=3)
  colnames(result)=c("X1","X2","X3")
  rownames(result)=c("Estimates","Variance")
  for (i in 1:n){
      x = runif(n)   # X1,...Xn~U(0,1)
      Est1[i]=mean(ftn(x))
      
      x=rexp(n,6)  # X1,...Xn~E(6)
      id=which((0<=x)&(x<=1))
      x=x[id]
      fx= 6*exp(-6*x)
      Est2[i]=mean(ftn(x)/fx)

      x=rbeta(n,1,2)   # X1,...Xn~Beta(1,2)
      Est3[i]=mean(ftn(x)/(2*(1-x)))
  }
   result[1,]=c(mean(Est1),mean(Est2),mean(Est3))
   result[2,]=c(var(Est1),var(Est2),var(Est3))
   return(result)
}
IS=ImportanceSampling(f,1000)
IS
save(IS,file="~/Documents/Cornell/Statistical computing/HW6/IS.rData")
```
```{r}
load(file="~/Documents/Cornell/Statistical computing/HW6/IS.rData")
IS
```


Cite: work with Rui Chen, rc687
