---
title: "writeup two"
author: "Tian Tan"
date: "May 10, 2016"
output: pdf_document
---

Problem 2

2a.Fit a simple linear regression model to the data. Display the fitted model on a scatterplot of the data.
```{r}
dataN=data.frame("nitrogen"= c(rep(0,4),rep(30,4),rep(60,4),rep(90,4),rep(120,4)),
           "yield"=c(1.41,1.75,2.02,2.13,1.93,2.24,2.29,2.35,2.12,2.38,
                     2.49,2.57,2.16,2.20,2.28,2.49,2.34,2.45,2.59,2.62))
model=lm(yield~nitrogen, dataN)
summary(model)
plot(dataN$nitrogen,dataN$yield,main="nitrogen & yield model",col="black")
lines(dataN$nitrogen,model$fitted.values,col="blue")
```

2b. Determine the form of the gradient vector of the model function $f(x,\theta) =\beta_0+\beta_1 min(x,N_{max})$ with respect to the parameter $\theta=\beta_0+\beta_1 min(x, N_{max})$.

So for $x<N_{max}$, $f(x,\theta)=\beta_0+\beta_1 x$, $\frac{df(x,\theta)}{d\theta_3} = 0$;

For $x>N_{max}$, $f(x,\theta)=\beta_0+\beta_1 \cdot N_{max}$, $\frac{df(x,\theta)}{d\theta_3}=\theta_2=\beta_1$;

So in conclusion, the gradient vector form is $\Delta f(x,\theta)=-(1, min(x,N_{max}),\beta_1 \cdot ((N>=N_{max})==1)$

2c. LPmodel function 
```{r}
LPmodel<-function(Y,N,tol=1e-5,maxiter=100){
  ## compute starting values inside the function
  ## del=change in sum of squares between iterations
  beta0=1.95550
  beta1=0.00475
  Nmax=max(N)-5
  theta=c(beta0,beta1,Nmax)
  
  fn<-function(t,x){
    # t is theta, x is the function variable
    return(t[1]+t[2]*pmin(x,t[3]))
  }
  dfn<-function(t,x){
    return(cbind(1,pmin(x,t[3]),t[2]*((x>=t[3])==TRUE)))
  }
  
  
  tol.met=FALSE
  iter=0
  del=0
  while(!tol.met){
    iter=iter+1
    oldtheta=theta
    f=fn(theta,N)
    g=dfn(theta,N)
    theta=theta+solve(t(g)%*%g,t(g)%*%(Y-f))
    fnew=fn(theta,N)
    del[iter]=sum((Y-f)^2)-sum((Y-fnew)^2)
    if (max(abs(theta-oldtheta))<tol|iter> maxiter){
      tol.met=TRUE
    }
  }
  return(list(b0=theta[1],b1=theta[2],Nmax=theta[3],iter=iter,del=del))
}
```

2d. Display the fitted model on the scatterplot. Use different colors for the simple linear regression and linear plateau model fits and indicate which is which using a legend.
```{r}
result=LPmodel(dataN$yield,dataN$nitrogen)
X=seq(1,120)
f=result$b0+result$b1*pmin(X,result$Nmax)
plot(dataN$nitrogen,dataN$yield,main="nitrogen & yield model",col="black")
lines(dataN$nitrogen,model$fitted.values,col="blue")
lines(X,f,col='red')
legend("bottomright",c("data","simple linear model","plateau models"),lty=c(1,1),lwd=c(2.5,2.5), col=c("BLACK","BlUE","RED"))
```









