---
title: "writeup six"
author: "Tian Tan"
date: "May 20, 2016"
output: pdf_document
---

6a. Given any square lower triangular matrix L and vector b, write code to find y such that Ly = b. Under what conditions can we not get a solution?
```{r}
findy<-function(L,b){
  # L is a square lower triangular matrix
  # b is a vector
  n=length(b)
  y=rep(0,n)
  y[1]=b[1]/L[1,1]
  for (i in 2:n){
    y[i]=b[i]
    for (j in 1:(i-1)){
      y[i]=y[i]-L[i,j]*y[j] 
    }
    y[i]=y[i]/L[i,i]
  }
  return(y)
}
```
The product of main entries of L cant't be 0. Moreover, if the length of b is not equal to the column number of L, then L and b can not be multiplied, thus we can't get a solution. 

6b. Using the same idea as the above, given any square upper triangular matrix U and vector y, write code to find x such that Ux = y.
```{r}
findx<-function(U,y){
  n=length(y)
  x=rep(0,n)
  x[n]=y[n]/U[n,n]
  for (i in (n-1):1){
    x[i]=y[i]
    for (j in n:(i+1)){
      x[i]=x[i]-U[i,j]*x[j] 
    }
    x[i]=x[i]/U[i,i]
  }
  return(x)
}
```

6c. Explain how you can use the code in parts 6a and 6b to find Ax = b given that you have the factorization A = LU. Thus, write a function:
```{r}
Findx<-function(A,b){
  #A is a n by n square matrix
  #b is a n by 1 column vector
  #x is a n by 1 column vectors such that A*x=b
  #You may use the package {\tt matrixcalc} or similar
  #but only to find the LU decomposition of a square matrix
  library(matrixcalc)
  LU_A=lu.decomposition(A)
  L=LU_A$L
  U=LU_A$U
  n=length(b)
  for (i in 1:n){
    if (L[i,i]==0 | U[i,i]==0 ){
      print("There is no solution")
      return(NULL)
    }
  }
  y=rep(0,n)
  y[1]=b[1]/L[1,1]
  for (i in 2:n){
    y[i]=b[i]
    for (j in 1:(i-1)){
      y[i]=y[i]-L[i,j]*y[j] 
    }
    y[i]=y[i]/L[i,i]
  }
  x=rep(0,n)
  x[n]=y[n]/U[n,n]
  for (i in (n-1):1){
    x[i]=y[i]
    for (j in n:(i+1)){
      x[i]=x[i]-U[i,j]*x[j] 
    }
    x[i]=x[i]/U[i,i]
  }
  return(x)
}
```

